\section{Experimental Setup: False-Negative Sensitivity (Observation~1)}
\label{app:false-negativity}

\paragraph{Data and prompt.}
We construct a synthetic summation task that gives us full control over which
context tokens are \emph{relevant} (i.e., must be attended to in order to
produce the correct answer). Each instance is a list of $N$ integers drawn
uniformly from $\{1,2,3\}$, formatted as the prompt
\textit{``Consider the list of numbers: $x_1, x_2, \ldots, x_N$. What is the
sum of the numbers? Answer with just one integer and nothing else. The answer
is''}. Greedy decoding produces the answer token, which is compared against
$\sum_i x_i$. We tokenise the prompt with the model's native tokenizer and
record, for each number $x_i$, the set of token positions that span its
surface form; these positions form the \emph{relevant} set $R$. Filler tokens
matching the closed-class set $\{$``the'', ``of'', ``a'', ``an'', ``and'',
``with'', ``to'', ``in'', ``on'', ``just'', ``one'', ``nothing'', ``else'',
``this'', ``,'', ``.'', ``:''$\}$ form the \emph{irrelevant} set $\bar R$.

\paragraph{Models and sparse-attention simulation.}
We evaluate Llama-3.2-1B/3B-Instruct, Llama-3.1-8B-Instruct, and
Qwen2.5-3B/7B-Instruct in \texttt{fp16} with eager attention. Sparse attention
is simulated by masking selected \emph{key} positions: we run a single
\texttt{forward} pass with a 4-D additive attention mask of shape
$(1,1,S,S)$ that combines the standard causal mask with $-\infty$ entries on
the columns of dropped positions, so that no query attends to those keys. This
is mathematically equivalent to evicting those KV-cache entries while keeping
all model weights and remaining keys/values intact. For Figure~\ref{fig:drop-error}
(Experiment~1) we drop $M{=}1$ token at a time, comparing one randomly chosen
token from $R$ with one randomly chosen token from $\bar R$, with $N{=}8$ and
$30$ random lists per model. For Figure~\ref{fig:error-spike} (Experiment~3)
we sweep $N\in\{2,3,4,5,6,8,10,12,16\}$ and a fixed-$K$ budget
$K\in\{2,3,4,6\}$ that retains exactly $K$ of the $N$ number tokens (chosen
uniformly at random) plus structural context (BOS, the question span, and all
non-list tokens), and compare against the dense baseline and an
oracle ``keep-all-relevant'' policy that retains every token in $R$. Each
configuration is averaged over $20$ random lists with seed $2$.

\paragraph{Metrics.}
We use two complementary metrics that together capture both decoding behaviour
and full distributional shift. (i) $P(\text{answer changed vs.\ dense})$ is
the fraction of trials in which greedy decoding under the dropped mask
produces a different output string than greedy decoding under dense attention,
when both are run for $4$ tokens; this is the metric reported in
Figure~\ref{fig:drop-error}. (ii) $\mathrm{KL}(p_{\text{drop}}\,\|\,p_{\text{dense}})$
is the Kullback--Leibler divergence between the next-token distributions
obtained from the dense and dropped forward passes after softmax over the full
vocabulary; this is the metric reported in Figure~\ref{fig:error-spike}. Both
are averaged across trials per $(N, \text{method})$ cell. Standard errors on
the bar plot use the binomial estimator
$\sqrt{p(1-p)/n}$.
